Learning Action Models from State Observations

### 2018-01-25 doc/ijcai-18/observations-learning.tex
updated sections 5 and 6 ijcai18 paper

\subsubsection{Evaluating with a test set}
When a reference model is not available, the evaluation of the learning process can be done with respect to a test set. Table~\ref{fig:observationstest} summarizes the obtained results when evaluating the quality of the learned models with respect to an observation test set. The table reports, for each domain, the {\em observation edit distance} (which is computed with our extended compilation), the {\em maximum edit distance}, and their ratio.

The reported results showed that, despite we are learning only from 25 state observations, twelve out of fifteen learned domains achieved ratios of $90\%$ or above. This fact evidences that the learned models required small edition amounts to match the observations given in the test set.

@@

The learning scores reported in Table~\ref{fig:observationsnomap} are poorer than in Table~\ref{fig:observationstest}. {\em Precision} and {\em recall},  given its syntax nature, report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model (changes of the roles of actions or their parameters).

@@

Table~\ref{fig:observationsmap} shows that significantly higher {\em precision} and {\em recall} scores are reported when comparing each learned action scheme $\xi\in\Xi$ with the corresponding reference scheme $f_{P\&R}(\xi)\in \Xi^*$ given by the $f_{P\&R}$ mapping. These results show that in all of the evaluated domains, except {\em ferry} and {\em satellite}, learning interchanged the roles of actions (or their parameters) with respect to their role in the reference model.

However in few domains, Table~\ref{fig:observationsmap} still reports lower scores than Table~\ref{fig:observationstest}. The explanation is that, how the observations of test sets are generated, affect to the results reported on Table~\ref{fig:observationstest}. For instance, in the {\em driverlog} domain the learned model missed learning the action scheme for the {\tt disembark-truck} action because this action was not considered in the learning set, the same happened in the {\em floortile} with the {\tt paint-down} action or in the {\em parking} domain with the {\tt move-curb-to-curb} action. We realized that these actions do not appear either in the corresponding test sets so the learned action models were not be penalized for not learning these action schemes.  The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.

\section{Conclusions}
As far as we know, this is the first work on learning \strips\ action models from state observations, exclusively using classical planning and evaluated over a wide range of different domains. Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented. ~\citeauthor{asai2017classical}~\citeyear{asai2017classical} proposed an approach for learning PDDL action models on a few domains from large images datasets (thousands of observations per domain).

The empirical results show that since our approach is strongly based on inference can generate good quaility models from very small data sets (ten out of fifteen domains achieved both {\em Precision} and {\em Recall} values over 0.75 learning only from 25 observations). Generating {\em informative} examples for learning planning action models is however an open issue. Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}. The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples. The combination of such exploration techniques with our learning approach is an intriguing research direction towards the bootstrapping of planning action models.

>>>>

\subsubsection{Evaluating with a test set}
When a reference model is not available, the evaluation of the learning process can be done with respect to a test set. Table~\ref{fig:observationstest} summarizes the obtained results when evaluating the quality of the learned models with respect to an observation test set. The table reports, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.

The reported results showed that, despite we are learning only from 25 state observations, twelve out of fifteen learned domains achieved ratios of $90\%$ or above. This fact evidences that the learned models required small edition amounts to match the state observations given in the test set.

@@

The learning scores reported in Table~\ref{fig:observationsnomap} are poorer than in Table~\ref{fig:observationstest}. {\em Precision} and {\em recall},  given its syntax nature, report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model (changes in the roles of the actions or their parameters).

@@

Table~\ref{fig:observationsmap} shows that significantly higher {\em precision} and {\em recall} scores are reported when comparing each learned action scheme $\xi\in\Xi$ with the corresponding reference scheme $f_{P\&R}(\xi)\in \Xi^*$ given by the $f_{P\&R}$ mapping. These results evidenced that in all of the evaluated domains, except {\em ferry} and {\em satellite}, learning interchanged the roles of actions or parameters with respect to their role in the reference model.

We can claim now that the action model for the {\em blocksworld} and the {\em gripper} domains was perfectly learned from only 25 state observations. However in few domains, Table~\ref{fig:observationsmap} still reports lower scores than Table~\ref{fig:observationstest}. The explanation is that, how the observations of test sets are generated, affect to the results reported on Table~\ref{fig:observationstest}. For instance, in the {\em driverlog} domain the learned model missed learning the action scheme for the {\tt disembark-truck} action because this action was not considered in the learning set. The same happened in the {\em floortile} with the {\tt paint-down} action and in {\em parking} with the {\tt move-curb-to-curb} action. We realized that these actions do not appear either in the corresponding test sets so the learned action models were not be penalized in the scores of Table~\ref{fig:observationstest} for not learning these action schemes.  %The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.

\section{Conclusions}
Unlike extensive-data ML approaches (like~\citeauthor{asai2017classical}~\citeyear{asai2017classical}), our work explores an alternative research direction to learn sound models from small amounts of state observations. As far as we know, this is the first work on learning \strips\ action models from state observations, exclusively using classical planning and evaluated over a wide range of different domains. Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.

The empirical results show that since our approach is strongly based on inference can generate good quality models from very small data sets. The action models of the {\em blocksworld} and the {\em gripper} were perfectly learned from only 25 observations besides the learned models, in ten out of fifteen domains, achieved both {\em Precision} and {\em Recall} values over 0.75.

Generating {\em informative} observations for learning planning action models is however an open issue. Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}. The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples. The combination of such exploration techniques with our learning approach is an intriguing research direction towards the bootstrapping of planning action models.
